---
title: "**Urban Air Quality And Health Impact Analysis - EDA Project - Team 2**"
author: "Abilasha Singh, Aswin Balaji Thippa Ramesh, Lixing Pan"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---


```{r setup, include=FALSE}
# Some of common RMD options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, message = F)
# Can globally set option for number display format.
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
```

## **Dataset** : Urban Air Quality and Health Impact Dataset

### **Variable** | **Definition** 

**DateTime** | Timestamp of the recorded data.

**City** | The U.S. city where the data was recorded (e.g., Phoenix, San Diego, New York City).

**Temp_Max** | Maximum temperature for the day (°F).

**Temp_Min** | Minimum temperature for the day (°F).

**Temp_Avg** | Average temperature for the day (°F).

**Feels_Like_Max** | Maximum "feels like" temperature for the day (°F).

**Feels_Like_Min** | Minimum "feels like" temperature for the day (°F).

**Feels_Like_Avg** | Average "feels like" temperature for the day (°F).

**Dew_Point** | Dew point temperature (°F).

**Humidity** | Relative humidity percentage.

**Precipitation** | Total precipitation for the day (inches).

**Precip_Prob** | Probability of precipitation (percentage).

**Precip_Cover** | Coverage of precipitation (percentage).

**Precip_Type** | Type of precipitation (e.g., rain, snow).

**Snow** | Amount of snowfall (inches).

**Snow_Depth** | Snow depth (inches).

**Wind_Gust** | Maximum wind gust speed (mph).

**Wind_Speed** | Average wind speed (mph).

**Wind_Direction** | Wind direction (degrees).

**Pressure** | Atmospheric pressure (hPa).

**Cloud_Cover** | Cloud cover percentage.

**Visibility** | Visibility distance (miles).

**Solar_Radiation** | Solar radiation (W/m²).

**Solar_Energy** | Solar energy received (kWh).

**UV_Index** | UV index level.

**Severe_Risk** | Risk level of severe weather (e.g., low, moderate, high).

**Sunrise** | Sunrise time (HH:MM:SS).

**Sunset** | Sunset time (HH:MM:SS).

**Moon_Phase** | Phase of the moon (e.g., new moon, full moon).

**Conditions** | General weather conditions (e.g., clear, cloudy).

**Description** | Detailed description of the weather conditions.

**Icon** | Weather icon representation.

**Stations** | Weather stations reporting data.

**Source** | Data source information.

**Temp_Range** | Temperature range for the day (difference between max and min temperatures).

**Heat_Index** | Heat index value for the day.

**Severity_Score** | Score representing the severity of weather conditions.

**Condition_Code** | Code representing specific weather conditions.

**Month** | Month of the year.

**Season** | Season of the year (e.g., winter, spring).

**Day_of_Wee**k | Day of the week.

**Is_Weekend** | Indicator if the day is a weekend.

**Health_Risk_Score** | Score representing the potential health risk based on weather and air quality conditions.

```{r}
library(dplyr)
library(ezids)
library(ggplot2)
library(tidyr)
library(reshape2)
library(gridExtra)
```

## **Data Loading**
```{r}
#loading the data
df=read.csv("air_quality_health_impact_data.csv")
head(df)
```

## **Data Description and Summmary**
```{r}
#Shape of the dataset
paste("Row Count:",dim(df)[1],"Column Count:",dim(df)[2])
```
```{r}
#comman stucture of the dataset
str(df)
```

```{r}
#summary of the dataset
summary(df)
```

```{r}
#Checking for unique values in all the columns
unique_counts <- sapply(df, function(x) length(unique(x)))
unique_counts
```