---
title: "**Urban Air Quality And Health Impact Analysis - EDA Project - Team 2**"
author: "Abilasha Singh, Aswin Balaji Thippa Ramesh, Lixing Pan"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---


```{r setup, include=FALSE}
# Some of common RMD options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, message = F)
# Can globally set option for number display format.
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
```

## **Dataset** : Urban Air Quality and Health Impact Dataset

### **Variable** | **Definition** 

**DateTime** | Timestamp of the recorded data.

**City** | The U.S. city where the data was recorded (e.g., Phoenix, San Diego, New York City).

**Temp_Max** | Maximum temperature for the day (°F).

**Temp_Min** | Minimum temperature for the day (°F).

**Temp_Avg** | Average temperature for the day (°F).

**Feels_Like_Max** | Maximum "feels like" temperature for the day (°F).

**Feels_Like_Min** | Minimum "feels like" temperature for the day (°F).

**Feels_Like_Avg** | Average "feels like" temperature for the day (°F).

**Dew_Point** | Dew point temperature (°F).

**Humidity** | Relative humidity percentage.

**Precipitation** | Total precipitation for the day (inches).

**Precip_Prob** | Probability of precipitation (percentage).

**Precip_Cover** | Coverage of precipitation (percentage).

**Precip_Type** | Type of precipitation (e.g., rain, snow).

**Snow** | Amount of snowfall (inches).

**Snow_Depth** | Snow depth (inches).

**Wind_Gust** | Maximum wind gust speed (mph).

**Wind_Speed** | Average wind speed (mph).

**Wind_Direction** | Wind direction (degrees).

**Pressure** | Atmospheric pressure (hPa).

**Cloud_Cover** | Cloud cover percentage.

**Visibility** | Visibility distance (miles).

**Solar_Radiation** | Solar radiation (W/m²).

**Solar_Energy** | Solar energy received (kWh).

**UV_Index** | UV index level.

**Severe_Risk** | Risk level of severe weather (e.g., low, moderate, high).

**Sunrise** | Sunrise time (HH:MM:SS).

**Sunset** | Sunset time (HH:MM:SS).

**Moon_Phase** | Phase of the moon (e.g., new moon, full moon).

**Conditions** | General weather conditions (e.g., clear, cloudy).

**Description** | Detailed description of the weather conditions.

**Icon** | Weather icon representation.

**Stations** | Weather stations reporting data.

**Source** | Data source information.

**Temp_Range** | Temperature range for the day (difference between max and min temperatures).

**Heat_Index** | Heat index value for the day.

**Severity_Score** | Score representing the severity of weather conditions.

**Condition_Code** | Code representing specific weather conditions.

**Month** | Month of the year.

**Season** | Season of the year (e.g., winter, spring).

**Day_of_Wee**k | Day of the week.

**Is_Weekend** | Indicator if the day is a weekend.

**Health_Risk_Score** | Score representing the potential health risk based on weather and air quality conditions.




```{r}
library(dplyr)
library(ezids)
library(ggplot2)
library(tidyr)
library(reshape2)
library(gridExtra)
```

**Libraries Used**

- **dplyr**: Facilitates data manipulation with functions for filtering, transforming, and summarizing data frames efficiently.
- **ezids**: Provides tools for generating, validating, and managing unique identifiers, essential for consistent data management.
- **ggplot2**: A powerful data visualization library that implements the Grammar of Graphics to create layered and customizable plots.
- **tidyr**: Simplifies the data tidying process with functions for reshaping data, including pivoting to wide and long formats.
- **reshape2**: Complements `tidyr` by offering tools for reshaping data between wide and long formats, ideal for complex data structures.
- **gridExtra**: Allows for arranging multiple `ggplot` objects and other grid-based plots on a single page, useful for comparisons.


## **Data Loading**
```{r}
#loading the data
df=read.csv("air_quality_health_impact_data.csv")
head(df)
```

## **Data Description and Summmary**
```{r}
#Shape of the dataset
paste("Row Count:",dim(df)[1],"Column Count:",dim(df)[2])
```

This indicates that the dataset contains **27,674** observations (rows) across **43** variables (columns), making it a substantial dataset for in-depth analysis.


```{r}
#Structure of the dataset
str(df)
```
**Dataset Structure**


Using the `str(df)` function, we inspect the dataset’s structure. This command provides a summary of each variable in the dataset, including:

- Data types (e.g., numeric, character).
- Sample values for each column.
- Overall dimensions of the dataset, confirming that it has 27,674 rows and 43 columns.

This inspection helps in understanding the types of data we’re working with, identifies any potential data cleaning needs, and sets up the foundation for the analysis by showing variable names and formats.


## **Appropriate Datatype Conversion**
```{r}
#Converting required columns as a factor variable
df$conditions <- as.factor(df$conditions)
df$description <- as.factor(df$description)
df$icon <- as.factor(df$icon)
df$source <- as.factor(df$source)
df$City <- as.factor(df$City)
df$Month <- as.factor(df$Month)
df$Season <- as.factor(df$Season)
df$Day_of_Week <- as.factor(df$Day_of_Week)
df$Is_Weekend <- as.factor(df$Is_Weekend)
```

```{r}
#Checking for datatype of every column
sapply(df, class)
```
To ensure accurate data processing, the code converts specific columns to the appropriate data types using the `as.factor()` function. This is essential for categorical variables, allowing for efficient analysis and visualization. The following columns are converted to factors: `conditions`, `description`, `icon`, `source`, `City`, `Month`, `Season`, `Day_of_Week`, and `Is_Weekend`. These variables are now treated as categorical data, which is important for statistical modeling and plotting.

After conversion, the `sapply(df, class)` function is used to check the data type of each column. This verification step confirms the conversions and provides an overview of all column data types:

- **Factor**: Categorical variables such as weather conditions, city, season, and day of the week.
- **Character**: Text-based columns like `datetime`, `sunrise`, and `sunset`.
- **Numeric**: Continuous variables such as temperature, humidity, and health risk scores.

This setup prepares the data for analysis by ensuring each variable is in a format suited to its type.


```{r}
#summary of the dataset
summary(df)
```

## **Dropping of unwanted columns**
```{r}
#Checking for unique values in all the columns
unique_counts <- sapply(df, function(x) length(unique(x)))
unique_counts
```

```{r}
#Dropping the unwanted columns
df<- subset(df, select = -c(Season,Month,snow,snowdepth)) 
head(df)
dim(df)
```

The code checks for the number of unique values in each column of the dataset to identify unnecessary columns. Based on this assessment, columns such as `Season`, `Month`, `snow`, and `snowdepth` are dropped from the dataframe to streamline the analysis. The updated dataframe is then displayed along with its dimensions.

## **Duplicates and missing values removal**
```{r}
#Checking for duplicate rows
duplicates <- df %>% duplicated() %>% sum()
print(paste("Number of duplicate rows:", duplicates))
```

```{r}
#Checking for missing values
missing_values <- colSums(is.na(df))
print(missing_values)
```
We found **no duplicate rows** in the dataset, confirming data integrity. Additionally, all columns were checked for missing values, and we confirmed that there are **zero missing values** across the dataset, indicating complete data availability.

## **Outliers Removal**
```{r}
library(ggplot2)
library(reshape2)

# Assuming df is your data frame
numeric_cols <- df %>% select_if(is.numeric)  # Select numeric columns

# Melt the data frame for ggplot2
melted_data <- melt(numeric_cols)

# Create combined boxplots using ggplot2
ggplot(melted_data, aes(x = variable, y = value)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  labs(title = "Combined Boxplots of Numeric Columns", 
       x = "Variables", 
       y = "Values") +
  theme_minimal() + # Optional: use a minimal theme for a cleaner look
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rot
```
```{r}
#Visualizing the outliers
numeric_cols <- df %>% select_if(is.numeric)
num_cols <- ncol(numeric_cols)

# Loop through numeric columns and create box plots
for (col in colnames(numeric_cols)) {
  boxplot(numeric_cols[[col]], 
          main = paste("Boxplot of", col), 
          ylab = col, 
          col = "lightblue", 
          border="darkblue")
}
```

```{r}
#Defining function to remove outliers using IQR approach
remove_outliers <- function(data) {
  for (col in names(data)) {
    if (is.numeric(data[[col]])) {
      Q1 <- quantile(data[[col]], 0.25)
      Q3 <- quantile(data[[col]], 0.75)
      IQR <- Q3 - Q1
      lower_bound <- Q1 - 1.5 * IQR
      upper_bound <- Q3 + 1.5 * IQR
      
      data <- data[data[[col]] >= lower_bound & data[[col]] <= upper_bound, ]
    }
  }
  return(data)
}
```

```{r}
#remove outliers by applying the function
df <- remove_outliers(df)
head(df)
```

```{r}
#Checking shape of the data after removing outliers
paste("Row Count:",dim(df)[1],"Column Count:",dim(df)[2])
```

We identified potential outliers using a box plot, which visualizes the distribution of the data and highlights values outside the whiskers. Outliers were defined as data points lying beyond 1.5 times the IQR above the third quartile or below the first quartile. These outliers were then removed to improve data quality for further analysis.


## **Univariate Analysis**
```{r}
# Distribution of Health Risk Score
ggplot(df, aes(x = Health_Risk_Score)) +
  geom_histogram(bins = 50, fill = "#2993ae", alpha = 0.5) +
  labs(title = "Distribution of Health Risk Score", x = "Health Risk Score", y = "Frequency")
```

```{r}
paste("The majority of the population has health risk scores around 9.0, with a small subset showing elevated scores near 10.0.")
```

## **Bivariate and Multivariate Analysis**

```{r}
paste("Is there a statistically significant difference in the Health Risk Score on weekends compared to weekdays?")
  
# Separate Health Risk Scores by weekends and weekdays
weekend_scores <- df$Health_Risk_Score[df$Is_Weekend == "True"]
weekday_scores <- df$Health_Risk_Score[df$Is_Weekend == "False"]

# Display hypothesis statements
cat("Hypothesis Statements:\n")
cat("Null Hypothesis (H0): There is no significant difference in the Health Risk Score between weekends and weekdays.\n")
cat("Alternative Hypothesis (H1): There is a significant difference in the Health Risk Score between weekends and weekdays.\n\n")

t_test_result <- t.test(weekend_scores, weekday_scores, alternative = "two.sided")

# Display the t-test results
print(t_test_result)

# Interpretation
cat("\nInterpretation:\n")
if(t_test_result$p.value < 0.05) {
  cat("Reject the null hypothesis: There is a statistically significant difference in the Health Risk Score between weekends and weekdays.\n")
} else {
  cat("Fail to reject the null hypothesis: There is no statistically significant difference in the Health Risk Score between weekends and weekdays.\n")
}
```
```{r}
boxplot(Health_Risk_Score ~ Is_Weekend, data = df,
        main = "Health Risk Score Comparison: Weekends vs. Weekdays",
        xlab = "Is Weekend",
        ylab = "Health Risk Score",
        col = c("lightblue", "lightgreen"),
        border = "#202020")
```

```{r}
# Boxplot of Health Risk Score by City
ggplot(df, aes(x = City, y = Health_Risk_Score, fill = City)) +
  geom_boxplot() +
  labs(title = "City-wise Health Risk Score", x = "City", y = "Health Risk Score") +
  theme(axis.text.x = element_text(angle =90,hjust=1))
```
**Key Observations**

- San Jose shows the widest IQR, indicating greater variability in health risk scores.
- Philadelphia has the lowest median health risk score, while New York City has the highest.
- Outliers are present in all cities, with New York City showing the most extreme values.

```{r}
# Boxplot of Windgust by City
ggplot(df, aes(x = City, y = windgust, fill = City)) +
  geom_boxplot() +
  labs(title = "City-wise windgust", x = "City", y = "Windgust") +
  theme(axis.text.x = element_text(angle =90,hjust=1))
```
**Key Observations**

- San Jose exhibits the widest IQR, indicating the most variability in wind gust speeds among the cities.
- Philadelphia has the lowest median wind gust speed, while Chicago has a relatively high median.
- New York City shows a few outliers, indicating instances of unusually high wind gust speeds.

```{r}
summary_stats <- df %>%
  group_by(City) %>%
  summarise(
    Avg_Dew_Point = mean(dew, na.rm = TRUE),
    Avg_Humidity = mean(humidity, na.rm = TRUE),
    Avg_Pressure = mean(pressure, na.rm = TRUE),
    Total_Precipitation = sum(precip, na.rm = TRUE),
    pressure_a = sum(pressure, na.rm = TRUE),
    .groups = 'drop'  # Drop grouping
  )

# Print summary statistics
print(summary_stats)

# Visualization

# 1. Average Dew Point by City
ggplot(summary_stats, aes(x = City, y = Avg_Dew_Point)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(title = "Average Dew Point by City", x = "City", y = "Average Dew Point") +
  theme_minimal()

# 2. Average Humidity by City
ggplot(summary_stats, aes(x = City, y = Avg_Humidity)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  labs(title = "Average Humidity by City", x = "City", y = "Average Humidity") +
  theme_minimal()

# 3. Total Precipitation by City
ggplot(summary_stats, aes(x = City, y = Total_Precipitation)) +
  geom_bar(stat = "identity", fill = "lightcoral") +
  labs(title = "Total Precipitation by City", x = "City", y = "Total Precipitation") +
  theme_minimal()

# 3. Pressue by City
ggplot(summary_stats, aes(x = City, y = pressure_a)) +
  geom_bar(stat = "identity", fill = "#909090") +
  labs(title = "Pressure by City", x = "City", y = "Pressure") +
  theme_minimal()
```

```{r}
#Corrrelation between every columns
constant_columns <- sapply(df, function(x) length(unique(x)) == 1)
df <- df[, !constant_columns] 

#correlation matrix
cor_matrix <- cor(df[, sapply(df, is.numeric)])

# Melt the correlation matrix for plotting
melted_cor_matrix <- melt(cor_matrix)

# Correlation heatmap
ggplot(melted_cor_matrix, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "green", high = "violet", mid = "white", midpoint = 0, limit = c(-1, 1)) +
  labs(title = "Correlation Heatmap", x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.text.y = element_text(vjust=0,hjust=0))
```
**Correlation with Health Risk Score**

A positive correlation means that as one variable increases, the other also increases. Conversely, a negative correlation means that as one variable increases, the other decreases.

Top **positive correlations:**

1. **Severity Score**: Strong positive link.
2. **Heat Index**: Significant positive association.
3. **Temperature Range**: Moderate positive correlation.
4. **Windgust**: Mild positive correlation.

Top **negative correlations:**

1. **Visibility**: Strong negative association.
2. **Solarenergy**: Significant negative correlation.
3. **Dew Point**: Moderate negative link.
4. **Humidity**: Mild negative correlation.

These indicate key factors influencing health risk scores.

```{r}
paste("Correlation between windgust and Health risk :",cor(df$windgust, df$Health_Risk_Score))
plot1<-ggplot(df, aes(x = windgust, y = Health_Risk_Score)) +
  geom_point(color = "#2993ae") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Wind Gusts vs Health Risk Score", 
       x = "Wind Gust", 
       y = "Health Risk Score") +
  theme_minimal()

# Calculate correlation
paste("Correlation between severity score and Health risk :",cor(df$Severity_Score, df$Health_Risk_Score))
plot2<-ggplot(df, aes(x = Severity_Score, y = Health_Risk_Score)) +
  geom_point(color = "#2993ae") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Severity Score vs Health Risk Score", 
       x = "Severity Score", 
       y = "Health Risk Score") +
  theme_minimal()

paste("Correlation between wind speed and Health risk :",cor(df$windspeed, df$Health_Risk_Score))
plot3<-ggplot(df, aes(x = windspeed, y = Health_Risk_Score)) +
  geom_point(color = "#2993ae") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Wind speed vs Health Risk Score", 
       x = "Wind Speed", 
       y = "Health Risk Score") +
  theme_minimal()

paste("Correlation between humidity and Health risk :",cor(df$humidity, df$Health_Risk_Score))
plot4<-ggplot(df, aes(x = humidity, y = Health_Risk_Score)) +
  geom_point(color = "#2993ae") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "humidity vs Health Risk Score", 
       x = "humidity", 
       y = "Health Risk Score") +
  theme_minimal()

paste("Correlation between humidity and windgust :",cor(df$humidity, df$Health_Risk_Score))
plot5<-ggplot(df, aes(x = humidity, y = windgust)) +
  geom_point(color = "violet") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "humidity vs windgust", 
       x = "humidity", 
       y = "windgust") +
  theme_minimal()

grid.arrange(plot1, plot2,plot3,plot4,plot5, ncol =2 )
```

The following analysis explores the relationships between environmental factors—such as wind gusts, severity score, wind speed, and humidity—and health risk scores.

1. **Direct Relationships with Health Risk Score**:
    - **Wind Gusts**: The scatter plot shows a slight positive trend, suggesting that higher wind gusts are associated with a marginal increase in health risk scores.
    - **Severity Score**: A strong positive correlation is observed, indicating that an increase in severity score corresponds to a notable rise in health risk scores.
    - **Wind Speed**: Wind speed also displays a positive relationship with health risk score, implying that as wind speed increases, so does the health risk.
    - **Humidity**: There is a positive association between humidity levels and health risk scores, with higher humidity associated with increased health risks.

2. **Indirect Influence of Humidity**:
    - **Humidity vs. Wind Gust**: The scatter plot shows a slight positive correlation, suggesting that higher humidity may contribute to increased wind gusts. Since wind gusts are directly linked with higher health risk scores, this indicates that humidity may indirectly influence health risks through its effect on wind conditions.
    
```{r}
paste("Correlation between visibility and Health risk :",cor(df$windgust, df$Severity_Score))
plot1<-ggplot(df, aes(x = visibility, y = Health_Risk_Score)) +
  geom_point(color = "#287743") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Visibility vs Health risk score", 
       x = "visibility", 
       y = "Health_Risk_Score") +
  theme_minimal()

# Calculate correlation
paste("Correlation between sunsetEpoch and Health risk :",cor(df$Severity_Score, df$Health_Risk_Score))
plot2<-ggplot(df, aes(x = sunsetEpoch, y = Health_Risk_Score)) +
  geom_point(color = "#287743") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Sunset Epoch vs Health Risk Score", 
       x = "Sunset Epoch", 
       y = "Health Risk Score") +
  theme_minimal()

paste("Correlation between sunriseEpoch and Health risk :",cor(df$windspeed, df$Health_Risk_Score))
plot3<-ggplot(df, aes(x = sunriseEpoch, y = Health_Risk_Score)) +
  geom_point(color = "#287743") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Sunrise Epoch vs Health Risk Score", 
       x = "Sunrise Epoch", 
       y = "Health Risk Score") +
  theme_minimal()

paste("Correlation between moonphase and Health risk :",cor(df$windspeed, df$Health_Risk_Score))
plot4<-ggplot(df, aes(x = moonphase, y = Health_Risk_Score)) +
  geom_point(color = "#287743") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Moon phase vs Health Risk Score", 
       x = "Moon phase", 
       y = "Health Risk Score") +
  theme_minimal()

grid.arrange(plot1, plot2,plot3,plot4, ncol = 2)
```
This analysis examines the relationship between certain environmental and astronomical factors—visibility, sunset epoch, sunrise epoch, and moon phase—and health risk scores.


- **Visibility**: The plot shows a slight negative correlation between visibility and health risk score. Higher visibility tends to be associated with lower health risks.
- **Sunset Epoch**: There is a negative trend between sunset time and health risk score, suggesting that later sunsets may be linked to reduced health risks.
- **Sunrise Epoch**: Similarly, a negative trend is observed between sunrise time and health risk score, with later sunrises associated with lower health risks.
- **Moon Phase**: The moon phase also shows a negative correlation with health risk score. As the moon progresses through its cycle, the health risk score tends to decrease.


These environmental and astronomical factors may subtly influence health risk scores. Clearer atmospheric conditions, later times of sunrise and sunset, and certain phases of the moon are associated with reduced health risks, suggesting possible connections between these natural phenomena and health outcomes.



<span style="color:navy;">Both health risk and wind gust scores fluctuate similarly over time, peaking around September 8th and declining sharply by mid- September. Health risk score is being influenced by windgust.</span>

```{r}
paste("How does the windgust vary over dates?")
  
# Convert datetime column to Date type (if not already done)
df$datetime <- as.Date(df$datetime, format = "%d-%m-%Y")  # Adjust the format as needed

# Extract month and day
df$Month <- format(df$datetime, "%m")  # Extract month as numeric
df$Day <- format(df$datetime, "%d")  # Extract day
# Average Health Risk Score by Date
daily_summary <- df %>%
  group_by(datetime) %>%
  summarise(Average_HRS = mean(windgust, na.rm = TRUE))

print(head(daily_summary))  # Preview the summary
library(ggplot2)

ggplot(daily_summary, aes(x = datetime, y = Average_HRS)) +
  geom_line( color = "blue") +
  geom_point(size = 1, color = "red") +
  labs(title = "Daily Windgust Score Over Time", 
       x = "Date", 
       y = "Windgust") +
  scale_x_date(date_breaks = "2 day", date_labels = "%d %b")
```


```{r}
paste("How does the Health Risk Score vary over dates?")
  
# Convert datetime column to Date type (if not already done)
df$datetime <- as.Date(df$datetime, format = "%d-%m-%Y")  # Adjust the format as needed

# Extract month and day
df$Month <- format(df$datetime, "%m")  # Extract month as numeric
df$Day <- format(df$datetime, "%d")  # Extract day
# Average Health Risk Score by Date
daily_summary <- df %>%
  group_by(datetime) %>%
  summarise(Average_HRS = mean(Health_Risk_Score, na.rm = TRUE))

print(head(daily_summary))  # Preview the summary
library(ggplot2)

ggplot(daily_summary, aes(x = datetime, y = Average_HRS)) +
  geom_line( color = "blue") +
  geom_point(size = 1, color = "red") +
  labs(title = "Daily Average Health Risk Score Over Time", 
       x = "Date", 
       y = "Average Health Risk Score") +
  scale_x_date(date_breaks = "2 day", date_labels = "%d %b")
```
Both graphs display a similar trend over the month of September. Initially, wind gusts are high, and health risk scores are elevated. As the month progresses, both the wind gust intensity and health risk scores decrease. By the end of the month, these values stabilize at lower levels. This refers to the correlation where reduced wind gusts are associated with lower health risks.


```{r}
daily_summary <- df %>%
  group_by(City, datetime) %>%  # Group by city and date
  summarise(Average_HRS = mean(Health_Risk_Score, na.rm = TRUE))

print(head(daily_summary))  # Preview the summary

# Load ggplot2 if not already loaded
library(ggplot2)

# Plot average Health Risk Score for each city
ggplot(daily_summary, aes(x = datetime, y = Average_HRS, color = City)) +
  geom_line() +  # Line for average scores
  geom_point(size = 1) +  # Points for average scores
  labs(title = "Health Risk Score Over Time by City", 
       x = "Date", 
       y = "Average Health Risk Score") +
  scale_x_date(date_breaks = "2 day", date_labels = "%d %b") +
  theme_minimal() +  # Use a minimal theme for better visualization
  theme(legend.position = "bottom")  # Place legend at the bottom
```



