---
title: "<span style='color:green; font-weight:bold;'>Urban Air Quality And Health Impact Analysis<br>   <span style='color:yellowgreen;font-size: 22pt'>EDA Project 1 - Team 2</span>"
author: "<span style='color:olive;'>Abilasha Singh, Aswin Balaji Thippa Ramesh, Lixing Pan</span><br><br>"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: false  
    toc_depth: 3
    toc_float: true
    theme: flatly
  pdf_document:
    toc: false 
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(scipen=9, digits = 6)
```
<span style="color:green; font-weight:bold; font-size: 20pt;">Research Summary</span><br>
<span style="color:navy;font-weight:bold;font-size: 13pt">1. Introduction</span><br>

<span style="color:navy;">Rising concerns about urban air quality and its impact on public health have prompted researchers to investigate the intricate links between environmental factors and health outcomes. This report provides a thorough analysis of how weather conditions relate to health risk scores in different cities during September. To achieve this, we began by developing several SMART questions focused on understanding how factors like temperature, humidity, and wind speed affect health risk scores. To reach our goals, we performed a detailed exploratory data analysis (EDA) to uncover patterns and trends in the dataset. We also used appropriate statistical tests to validate our findings and draw meaningful conclusions. This report aims to clarify the connections between environmental conditions and health.</span><br>

<span style="color:navy;font-weight:bold;font-size: 13pt">2. Dataset Summary</span><br>

<span style="color:navy;font-weight:bold;font-size: 11pt">2.1  Dataset Overview</span><br>

<span style="color:navy;">The dataset comprises a collection of 27,674 observations related to meteorological and environmental data, including temperature, humidity, and wind speed, recorded across across seven major U.S. cities for the month of September. It contains a total of 43 variables, such as Health Risk Score and Severity Score, providing a detailed view of daily atmospheric conditions.Key variables such as temperature, humidity, wind speed, visibility, and weather descriptions enable an in-depth analysis of how urban air quality affects health.</span>

```{r,results='hide'}
library(dplyr)
library(ezids)
library(ggplot2)
library(tidyr)
library(reshape2)
library(gridExtra)
```
<span style="color:navy;font-weight:bold;font-size: 10pt">Data Loading</span><br>

```{r}
#loading the data
df=read.csv("air_quality_health_impact_data.csv")
head(df)
```
<span style="color:navy;font-weight:bold;font-size: 10pt">Data Description and Summmary</span><br>
```{r}
#Shape of the dataset
paste("Row Count:",dim(df)[1],"Column Count:",dim(df)[2])
```

```{r}
#Structure of the dataset
str(df)
```
<span style="color:navy;font-weight:bold;font-size: 11pt">2.2  Data Source</span><br>

<span style="color:navy;">The dataset is presumed to originate from a weather forecasting or environmental aggregation service and is sourced from the Kaggle website.</span>

<span style="color:navy;font-weight:bold;font-size: 11pt">2.3  Scope and SMART Questions</span><br>

<span style="color:navy;">Key variables such as temperature, humidity, wind speed, visibility, and weather descriptions facilitate a comprehensive analysis of the effects of urban air quality on health. To guide our investigation, we have formulated several SMART questions:</span><br>
<span style="color:navy;">
1.How health risk scores differ between weekdays and weekends?<br>
2.Analyzing how key meteorological factors vary across different cities.<br>
3.Investigating the key meteorological factors that significantly influence health risk scores?<br>
4.How changes in humidity might affect overall health?<br>
5.Examining how wind speed and health risk scores vary over days.</span><br>

<span style="color:navy;">These questions will help us effectively explore the connections between environmental conditions and health outcomes.
</span><br>

<span style="color:navy;font-weight:bold;font-size: 13pt">3. Limitations of the Dataset</span><br>

<span style="color:navy;">The dataset has temporal and spatial limitations, as it only contains data from September and a limited number of cities. This may not adequately represent the weather patterns over a broader range of years or the associated health risks. Furthermore, it lacks several important variables that could clarify how health risks were calculated, potentially overlooking significant factors that impact health.</span>

<span style="color:navy;font-weight:bold;font-size: 13pt">4. Exploratory Data Analysis</span><br>

<span style="color:navy;font-weight:bold;font-size: 11pt">4.1 Appropriate Datatype Conversion</span><br>

<span style="color:navy;">Out of 43 variables, 9 variables are factor, 30 are numeric, and 3 are character.</span>

```{r}
#Converting required columns as a factor variable
df$conditions <- as.factor(df$conditions)
df$description <- as.factor(df$description)
df$icon <- as.factor(df$icon)
df$source <- as.factor(df$source)
df$City <- as.factor(df$City)
df$Month <- as.factor(df$Month)
df$Season <- as.factor(df$Season)
df$Day_of_Week <- as.factor(df$Day_of_Week)
df$Is_Weekend <- as.factor(df$Is_Weekend)
```

```{r}
#Checking for datatype of every column
sapply(df, class)
```

```{r}
#summary of the dataset
summary(df)
```

<span style="color:navy;font-weight:bold;font-size: 11pt">4.2 Dropping unwanted columns</span><br>

<span style="color:navy;">Dropping season , snow, snowdepth and month columns as they have single value.</span>

```{r}
#Checking for unique values in all the columns
unique_counts <- sapply(df, function(x) length(unique(x)))
unique_counts
```

```{r}
#Dropping the unwanted columns
df<- subset(df, select = -c(Season,Month,snow,snowdepth)) 
head(df)
dim(df)
```

<span style="color:navy;font-weight:bold;font-size: 11pt">4.3 Duplicates and missing values removal</span><br>

<span style="color:navy;">No missing values and duplicate rows</span>

```{r}
#Checking for duplicate rows
duplicates <- df %>% duplicated() %>% sum()
print(paste("Number of duplicate rows:", duplicates))
```

```{r}
#Checking for missing values
missing_values <- colSums(is.na(df))
print(missing_values)
```

<span style="color:navy;font-weight:bold;font-size: 11pt">4.4 Outliers Removal</span><br>

<span style="color:navy;">After removing the outliers, we are left with 18885 observations
</span>

```{r}
#Visualizing the outliers
numeric_cols <- df %>% select_if(is.numeric)
num_cols <- ncol(numeric_cols)

# Loop through numeric columns and create box plots
for (col in colnames(numeric_cols)) {
  boxplot(numeric_cols[[col]], 
          main = paste("Boxplot of", col), 
          ylab = col, 
          col = "lightblue", 
          border="darkblue")
}
```

```{r}
#Defining function to remove outliers using IQR approach
remove_outliers <- function(data) {
  for (col in names(data)) {
    if (is.numeric(data[[col]])) {
      Q1 <- quantile(data[[col]], 0.25)
      Q3 <- quantile(data[[col]], 0.75)
      IQR <- Q3 - Q1
      lower_bound <- Q1 - 1.5 * IQR
      upper_bound <- Q3 + 1.5 * IQR
      
      data <- data[data[[col]] >= lower_bound & data[[col]] <= upper_bound, ]
    }
  }
  return(data)
}
```

```{r}
#remove outliers by applying the function
df <- remove_outliers(df)
head(df)
```

```{r}
#Checking shape of the data after removing outliers
paste("Row Count:",dim(df)[1],"Column Count:",dim(df)[2])
```

<span style="color:navy;font-weight:bold;font-size: 11pt">4.5 Univariate Analysis</span><br>

<span style="color:navy;">Most of the population has health risk scores around 9.0, with a small subset showing elevated scores near 10.0.</span>

```{r}
# Distribution of Health Risk Score
ggplot(df, aes(x = Health_Risk_Score)) +
  geom_histogram(bins = 50, fill = "#2993ae", alpha = 0.5) +
  labs(title = "Distribution of Health Risk Score", x = "Health Risk Score", y = "Frequency")
```

```{r}
paste("The majority of the population has health risk scores around 9.0, with a small subset showing elevated scores near 10.0.")
```

<span style="color:navy;font-weight:bold;font-size: 11pt">4.6 Bivariate and Multivariate Analysis</span><br>

<span style="color:navy;">There is a significant difference in Health Risk Scores between weekends and weekdays. Scores are generally higher on weekends than on weekdays, indicating increased health risks during weekends</span>

```{r}
paste("Is there a statistically significant difference in the Health Risk Score on weekends compared to weekdays?")
  
# Separate Health Risk Scores by weekends and weekdays
weekend_scores <- df$Health_Risk_Score[df$Is_Weekend == "True"]
weekday_scores <- df$Health_Risk_Score[df$Is_Weekend == "False"]

# Display hypothesis statements
cat("Hypothesis Statements:\n")
cat("Null Hypothesis (H0): There is no significant difference in the Health Risk Score between weekends and weekdays.\n")
cat("Alternative Hypothesis (H1): There is a significant difference in the Health Risk Score between weekends and weekdays.\n\n")

t_test_result <- t.test(weekend_scores, weekday_scores, alternative = "two.sided")

# Display the t-test results
print(t_test_result)

# Interpretation
cat("\nInterpretation:\n")
if(t_test_result$p.value < 0.05) {
  cat("Reject the null hypothesis: There is a statistically significant difference in the Health Risk Score between weekends and weekdays.\n")
} else {
  cat("Fail to reject the null hypothesis: There is no statistically significant difference in the Health Risk Score between weekends and weekdays.\n")
}
```
```{r}
boxplot(Health_Risk_Score ~ Is_Weekend, data = df,
        main = "Health Risk Score Comparison: Weekends vs. Weekdays",
        xlab = "Is Weekend",
        ylab = "Health Risk Score",
        col = c("lightblue", "lightgreen"),
        border = "#202020")
```

```{r}
paste("The weak negative correlation of -0.24 suggests that higher temperatures are slightly associated with lower health risk scores, though the relationship is not strong")
```

<span style="color:navy;">San Jose has the highest health risk score and wind gusts, while Philadelphia shows the lowest health risk score and Los Angeles the lowest wind gusts.</span>

```{r}
# Boxplot of Health Risk Score by City
ggplot(df, aes(x = City, y = Health_Risk_Score, fill = City)) +
  geom_boxplot() +
  labs(title = "City-wise Health Risk Score", x = "City", y = "Health Risk Score") +
  theme(axis.text.x = element_text(angle =90,hjust=1))
```

```{r}
# Boxplot of Windgust by City
ggplot(df, aes(x = City, y = windgust, fill = City)) +
  geom_boxplot() +
  labs(title = "City-wise windgust", x = "City", y = "Windgust") +
  theme(axis.text.x = element_text(angle =90,hjust=1))
```

<span style="color:navy;">Visualize the strength and direction of relationships (correlations) between pairs of variables in a dataset.</span>

```{r}
#Correlation between every columns
constant_columns <- sapply(df, function(x) length(unique(x)) == 1)
df <- df[, !constant_columns] 

#correlation matrix
cor_matrix <- cor(df[, sapply(df, is.numeric)])

# Melt the correlation matrix for plotting
melted_cor_matrix <- melt(cor_matrix)

# Correlation heatmap
ggplot(melted_cor_matrix, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "green", high = "violet", mid = "white", midpoint = 0, limit = c(-1, 1)) +
  labs(title = "Correlation Heatmap", x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.text.y = element_text(vjust=0,hjust=0))
```

<span style="color:navy;">Wind gusts have a positive impact on health risk, whereas temperature has a weaker, slightly negative relationship with health risk scores.Wind-related factors (gusts and speed) and humidity increase health risk scores, while factors like visibility, sunrise/sunset times, and moon phases reduce them.</span>

```{r}
paste("Does wind gust have a significant impact on the Health Risk Score?")

paste("Correlation between windgust and Health risk :",cor(df$windgust, df$Health_Risk_Score))
plot1<-ggplot(df, aes(x = windgust, y = Health_Risk_Score)) +
  geom_point(color = "#2993ae") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Wind Gusts vs Health Risk Score", 
       x = "Wind Gust", 
       y = "Health Risk Score") +
  theme_minimal()

# Calculate correlation
paste("Correlation between severity score and Health risk :",cor(df$Severity_Score, df$Health_Risk_Score))
plot2<-ggplot(df, aes(x = Severity_Score, y = Health_Risk_Score)) +
  geom_point(color = "#2993ae") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Severity Score vs Health Risk Score", 
       x = "Severity Score", 
       y = "Health Risk Score") +
  theme_minimal()

paste("Correlation between wind speed and Health risk :",cor(df$windspeed, df$Health_Risk_Score))
plot3<-ggplot(df, aes(x = windspeed, y = Health_Risk_Score)) +
  geom_point(color = "#2993ae") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Wind speed vs Health Risk Score", 
       x = "Wind Speed", 
       y = "Health Risk Score") +
  theme_minimal()

paste("Correlation between humidity and Health risk :",cor(df$humidity, df$Health_Risk_Score))
plot4<-ggplot(df, aes(x = humidity, y = Health_Risk_Score)) +
  geom_point(color = "#2993ae") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "humidity vs Health Risk Score", 
       x = "humidity", 
       y = "Health Risk Score") +
  theme_minimal()

paste("Correlation between humidity and windgust :",cor(df$humidity, df$Health_Risk_Score))
plot5<-ggplot(df, aes(x = humidity, y = windgust)) +
  geom_point(color = "violet") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "humidity vs windgust", 
       x = "humidity", 
       y = "windgust") +
  theme_minimal()

grid.arrange(plot1, plot2,plot3,plot4,plot5, ncol =2 )
```

```{r}
paste("Correlation between visibility and Health risk :",cor(df$windgust, df$Severity_Score))
plot1<-ggplot(df, aes(x = visibility, y = Health_Risk_Score)) +
  geom_point(color = "#287743") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Visibility vs Health risk score", 
       x = "visibility", 
       y = "Health_Risk_Score") +
  theme_minimal()

# Calculate correlation
paste("Correlation between sunsetEpoch and Health risk :",cor(df$Severity_Score, df$Health_Risk_Score))
plot2<-ggplot(df, aes(x = sunsetEpoch, y = Health_Risk_Score)) +
  geom_point(color = "#287743") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Sunset Epoch vs Health Risk Score", 
       x = "Sunset Epoch", 
       y = "Health Risk Score") +
  theme_minimal()

paste("Correlation between sunriseEpoch and Health risk :",cor(df$windspeed, df$Health_Risk_Score))
plot3<-ggplot(df, aes(x = sunriseEpoch, y = Health_Risk_Score)) +
  geom_point(color = "#287743") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Sunrise Epoch vs Health Risk Score", 
       x = "Sunrise Epoch", 
       y = "Health Risk Score") +
  theme_minimal()

paste("Correlation between moonphase and Health risk :",cor(df$windspeed, df$Health_Risk_Score))
plot4<-ggplot(df, aes(x = moonphase, y = Health_Risk_Score)) +
  geom_point(color = "#287743") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Moon phase vs Health Risk Score", 
       x = "Moon phase", 
       y = "Health Risk Score") +
  theme_minimal()

grid.arrange(plot1, plot2,plot3,plot4, ncol = 2)
```
```{r}
summary_stats <- df %>%
  group_by(City) %>%
  summarise(
    Avg_Dew_Point = mean(dew, na.rm = TRUE),
    Avg_Humidity = mean(humidity, na.rm = TRUE),
    Avg_Pressure = mean(pressure, na.rm = TRUE),
    Total_Precipitation = sum(precip, na.rm = TRUE),
    pressure_a = sum(pressure, na.rm = TRUE),
    .groups = 'drop'  # Drop grouping
  )

# Print summary statistics
print(summary_stats)

# Visualization

# 1. Average Dew Point by City
ggplot(summary_stats, aes(x = City, y = Avg_Dew_Point)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(title = "Average Dew Point by City", x = "City", y = "Average Dew Point") +
  theme_minimal()

# 2. Average Humidity by City
ggplot(summary_stats, aes(x = City, y = Avg_Humidity)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  labs(title = "Average Humidity by City", x = "City", y = "Average Humidity") +
  theme_minimal()

# 3. Total Precipitation by City
ggplot(summary_stats, aes(x = City, y = Total_Precipitation)) +
  geom_bar(stat = "identity", fill = "lightcoral") +
  labs(title = "Total Precipitation by City", x = "City", y = "Total Precipitation") +
  theme_minimal()

# 3. Pressue by City
ggplot(summary_stats, aes(x = City, y = pressure_a)) +
  geom_bar(stat = "identity", fill = "#909090") +
  labs(title = "Pressure by City", x = "City", y = "Pressure") +
  theme_minimal()
```

<span style="color:navy;">Both health risk and wind gust scores fluctuate similarly over time, peaking around September 8th and declining sharply by mid- September. Health risk score is being influenced by windgust.</span>

```{r}
paste("How does the windgust vary over dates?")
  
# Convert datetime column to Date type (if not already done)
df$datetime <- as.Date(df$datetime, format = "%d-%m-%Y")  # Adjust the format as needed

# Extract month and day
df$Month <- format(df$datetime, "%m")  # Extract month as numeric
df$Day <- format(df$datetime, "%d")  # Extract day
# Average Health Risk Score by Date
daily_summary <- df %>%
  group_by(datetime) %>%
  summarise(Average_HRS = mean(windgust, na.rm = TRUE))

print(head(daily_summary))  # Preview the summary
library(ggplot2)

ggplot(daily_summary, aes(x = datetime, y = Average_HRS)) +
  geom_line( color = "blue") +
  geom_point(size = 1, color = "red") +
  labs(title = "Daily Windgust Score Over Time", 
       x = "Date", 
       y = "Windgust") +
  scale_x_date(date_breaks = "2 day", date_labels = "%d %b")
```
```{r}
paste("How does the Health Risk Score vary over dates?")
  
# Convert datetime column to Date type (if not already done)
df$datetime <- as.Date(df$datetime, format = "%d-%m-%Y")  # Adjust the format as needed

# Extract month and day
df$Month <- format(df$datetime, "%m")  # Extract month as numeric
df$Day <- format(df$datetime, "%d")  # Extract day
# Average Health Risk Score by Date
daily_summary <- df %>%
  group_by(datetime) %>%
  summarise(Average_HRS = mean(Health_Risk_Score, na.rm = TRUE))

print(head(daily_summary))  # Preview the summary
library(ggplot2)

ggplot(daily_summary, aes(x = datetime, y = Average_HRS)) +
  geom_line( color = "blue") +
  geom_point(size = 1, color = "red") +
  labs(title = "Daily Average Health Risk Score Over Time", 
       x = "Date", 
       y = "Average Health Risk Score") +
  scale_x_date(date_breaks = "2 day", date_labels = "%d %b")
```
```{r}
daily_summary <- df %>%
  group_by(City, datetime) %>%  # Group by city and date
  summarise(Average_HRS = mean(Health_Risk_Score, na.rm = TRUE))

print(head(daily_summary))  # Preview the summary

# Load ggplot2 if not already loaded
library(ggplot2)

# Plot average Health Risk Score for each city
ggplot(daily_summary, aes(x = datetime, y = Average_HRS, color = City)) +
  geom_line() +  # Line for average scores
  geom_point(size = 1) +  # Points for average scores
  labs(title = "Health Risk Score Over Time by City", 
       x = "Date", 
       y = "Average Health Risk Score") +
  scale_x_date(date_breaks = "2 day", date_labels = "%d %b") +
  theme_minimal() +  # Use a minimal theme for better visualization
  theme(legend.position = "bottom")  # Place legend at the bottom
```

<span style="color:navy;font-weight:bold;font-size: 13pt">5. Conclusion and EDA Insights</span><br>

<span style="color:navy;">Health Risk Variation: Health risk scores are higher on weekends compared to weekdays, indicating greater risks during weekends.</span>

<span style="color:navy;">City Comparison: San Jose has the highest health risk and wind gusts, while Philadelphia shows the lowest health risk.</span>

<span style="color:navy;">Meteorological Impact: Wind gusts significantly increase health risk scores, while temperature has a weaker influence.</span>

<span style="color:navy;">Humidity and Health: Health risk scores fluctuate with changes in wind gusts, peaking in early September.</span>

<span style="color:navy;">Key Factors: Wind-related factors increase health risks, while visibility and sunrise/sunset times lowers them indirectly.
</span>

<span style="color:navy;font-weight:bold;font-size: 13pt">6. Further Research and Limitations</span><br>

<span style="color:navy;font-weight:bold;font-size: 13pt">7. References</span><br><br>








