---
title: "URBAN AIR QUALITY AND HEALTH IMPACT ANALYSIS"
author: "Project Team: Team 2 (Aswin Balaji Thippa Ramesh, Abhilasha Singh, Lixing Pan)"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=F}
# The package "ezids" (EZ Intro to Data Science) includes some helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
# You will need to install it (once) from GitHub.
# library(devtools)
# devtools::install_github("physicsland/ezids")
# Then load the package in your R session.
library(ezids)
```


```{r setup, include=FALSE}
# Some of common RMD options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, message = F)
# Can globally set option for number display format.
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
```

```{r}
# 1. Do not provide answers/comments inside code blocks (like here) -- those are notes between coders/self and will be ignored for grading. 
# 2. Make sure your knitr options are set to include all results/code to be graded in the final document.
# 3. All charts/graphs/tables should have appropriate titles/labels/captions. 
# 4. Compose your answers using inline R code instead of using the code-block output as much as you can. 
# 5. Your grade is also determined by the style. Even if you answer everything correctly, but the .html does not look appealing, you will not get full credit. Pay attention to the details that we mentioned in class/homework and in previous sample .Rmd files. For example, how to use #, ##, ###, ..., bold face, italics, inline codes, tables, ..., {results = "asis"}, use of colors in plots/ggplots, and so forth.
```

The write-up should also provide
thoughtful insight into the process of the analysis, including its strengths and weaknesses.
Potential area of topics to address in this summary may include:
   What do we know about this dataset?
   What are the limitations of the dataset?
   How was the information gathered?
   What analysis has already been completed related to the content in your dataset?
   How did the research you gathered contribute to your question development?
   What additional information would be beneficial?
   How did your question change, if at all, after Exploratory Data Analysis?
   Based on EDA can you begin to sketch out an answer to your question?
   References (APA style preferred)
  
```{r}
Data <- read.csv("C:/Abhilasha/Introduction to Data Science(6101)/Project1/Urban Air Quality and Health Impact Dataset.csv", header = TRUE)
str(Data)
summary(Data)
```


```{r}
```

Data Cleaning:

Missing values in the dataset

```{r}
missing_summary <- sapply(Data, function(x) sum(is.na(x)))
print(missing_summary)
Data_WMV <- na.omit(Data)
str(Data_WMV)
```

```{r}
library(corrplot)
corr_matrix <- cor(Data[, sapply(Data, is.numeric)], use = "complete.obs")

# Plot the correlation matrix
corrplot(corr_matrix, method = "circle")
```

missing values in the dataset

```{r}
missing_summary <- sapply(Data, function(x) sum(is.na(x)))
print(missing_summary)
```

Remove Rows with Missing Data

```{r}
```