---
title: "**Urban Air Quality And Health Impact Analysis**"
author: "Abilasha Singh, Aswin Balaji Thippa Ramesh, Lixing Pan"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F)
options(scientific=T, digits = 3) 
```
# **Introduction**

Rising concerns about urban air quality and its impact on public health have prompted researchers to investigate the intricate links between environmental factors and health outcomes. This report provides a thorough analysis of how weather conditions relate to health risk scores in different cities during September. To achieve this, we began by developing several SMART questions focused on understanding how factors like temperature, humidity, and wind speed affect health risk scores. To reach our goals, we performed a detailed exploratory data analysis (EDA) to uncover patterns and trends in the dataset. We also used appropriate statistical tests to validate our findings and draw meaningful conclusions. This report aims to clarify the connections between environmental conditions and health

# **Dataset Summary**

### a) Data Source

The dataset is presumed to originate from a weather forecasting or environmental aggregation service and is sourced from the Kaggle website.

### b) Dataset Limitations

The dataset has temporal and spatial limitations, as it only contains data from September and a limited number of cities. This may not adequately represent the weather patterns over a broader range of years or the associated health risks. Furthermore, it lacks several important variables that could clarify how health risks were calculated, potentially overlooking significant factors that impact health.

# **Scope and SMART Questions**

Key variables such as `temperature`, `humidity`, `wind speed`, `visibility`, and `weather descriptions` facilitate a comprehensive analysis of the effects of urban air quality on health. 

To guide our investigation, we have formulated several SMART questions:

1. How do health risk scores differ between weekdays and weekends?

2. How do key meteorological factors vary across different cities?

3. Which key meteorological factors significantly influence health risk scores?

4. How might changes in humidity affect overall health?

5. How do wind speed and health risk scores vary over time?

These questions will help us effectively explore the connections between environmental conditions and health outcomes.


```{r,include = FALSE}
library(dplyr)
library(ezids)
library(ggplot2)
library(tidyr)
library(reshape2)
library(gridExtra)
```
# **Data Loading and Data Description**

### a) Data Loading
```{r}
#loading the data
df=read.csv("air_quality_health_impact_data.csv")
head(df)
```


### b) Shape of the Dataset

```{r}
#Shape of the dataset
paste("Row Count:",dim(df)[1],"Column Count:",dim(df)[2])
```

The dataset comprises a collection of **27,674** observations related to meteorological and environmental data, including temperature, humidity, and wind speed, recorded across across seven major U.S. cities for the month of September. It contains a total of **43** variables, such as `Health Risk Score` and `Severity Score`, providing a detailed view of daily atmospheric conditions.Key variables such as `temperature`, `humidity`, `wind speed`, `visibility`, and `weather descriptions` enable an in-depth analysis of how urban air quality affects health.</span>

### c) Structure of the Dataset
```{r}
#Structure of the dataset
str(df)
```
```{r}
#summary of the dataset
summary(df)
```


We inspect the dataset’s structure and it provides a summary of each variable in the dataset, including:

- Data types (e.g., numeric, character).
- Sample values for each column.
- Overall dimensions of the dataset, confirming that it has 27,674 rows and 43 columns.

This inspection helps in understanding the types of data we’re working with, identifies any potential data cleaning needs, and sets up the foundation for the analysis by showing variable names and formats.


# **Data Cleaning**

### a) Appropriate Datatype Conversion

To ensure accurate data processing, we convert specific columns to the appropriate data types using the `as.factor()` function. This is essential for categorical variables, allowing for efficient analysis and visualization. The following columns are converted to factors: `conditions`, `description`, `icon`, `source`, `City`, `Month`, `Season`, `Day_of_Week`, and `Is_Weekend`. These variables are now treated as categorical data, which is important for statistical modeling and plotting.

```{r}
#Converting required columns as a factor variable
df$conditions <- as.factor(df$conditions)
df$description <- as.factor(df$description)
df$icon <- as.factor(df$icon)
df$source <- as.factor(df$source)
df$City <- as.factor(df$City)
df$Month <- as.factor(df$Month)
df$Season <- as.factor(df$Season)
df$Day_of_Week <- as.factor(df$Day_of_Week)
df$Is_Weekend <- as.factor(df$Is_Weekend)
```

```{r}
#Checking for datatype of every column
sapply(df, class)
```

After conversion, we check the data type of each column. This verification step confirms the conversions and provides an overview of all column data types:

- **Factor**: Categorical variables such as weather conditions, city, season, and day of the week.
- **Character**: Text-based columns like `datetime`, `sunrise`, and `sunset`.
- **Numeric**: Continuous variables such as temperature, humidity, and health risk scores.

Out of 43 variables, 9 variables are factor, 30 are numeric, and 3 are character.

This setup prepares the data for analysis by ensuring each variable is in a format suited to its type.

```{r}
#summary of the dataset
summary(df)
```

<span style="color:navy;"></span>

### b) Dropping Unwanted Columns

```{r}
#Checking for unique values in all the columns
unique_counts <- sapply(df, function(x) length(unique(x)))
unique_counts
```

```{r}
#Dropping the unwanted columns
df<- subset(df, select = -c(Season,Month,snow,snowdepth)) 
head(df)
dim(df)
```

We check for the number of unique values in each column of the dataset to identify unnecessary columns. Based on this assessment, columns such as `Season`, `Month`, `snow`, and `snowdepth` are dropped from the dataframe as they have single value to streamline the analysis. The updated dataframe is then displayed along with its dimensions. 

### c) Duplicates and missing values removal

```{r}
#Checking for duplicate rows
duplicates <- df %>% duplicated() %>% sum()
print(paste("Number of duplicate rows:", duplicates))
```

```{r}
#Checking for missing values
missing_values <- colSums(is.na(df))
print(missing_values)
```

We found **no duplicate rows** in the dataset, confirming data integrity. Additionally, all columns were checked for missing values, and we confirmed that there are **zero missing values** across the dataset, indicating complete data availability.

### d) Outliers Removal

```{r}
#Visualizing the outliers
numeric_cols <- df %>% select_if(is.numeric)
num_cols <- ncol(numeric_cols)

# Loop through numeric columns and create box plots
for (col in colnames(numeric_cols)) {
  boxplot(numeric_cols[[col]], 
          main = paste("Boxplot of", col), 
          ylab = col, 
          col = "lightblue", 
          border="darkblue")
}
```

```{r}
#Defining function to remove outliers using IQR approach
remove_outliers <- function(data) {
  for (col in names(data)) {
    if (is.numeric(data[[col]])) {
      Q1 <- quantile(data[[col]], 0.25)
      Q3 <- quantile(data[[col]], 0.75)
      IQR <- Q3 - Q1
      lower_bound <- Q1 - 1.5 * IQR
      upper_bound <- Q3 + 1.5 * IQR
      
      data <- data[data[[col]] >= lower_bound & data[[col]] <= upper_bound, ]
    }
  }
  return(data)
}
```

```{r}
#remove outliers by applying the function
df <- remove_outliers(df)
head(df)
```

```{r}
#Checking shape of the data after removing outliers
paste("Row Count:",dim(df)[1],"Column Count:",dim(df)[2])
```

We identified potential outliers using a box plot, which visualizes the distribution of the data and highlights values outside the whiskers. Outliers were defined as data points lying beyond 1.5 times the IQR above the third quartile or below the first quartile. These outliers were then removed to improve data quality for further analysis.After removing the outliers, we are left with 18885 observations

# **Exploratory Data Analysis**

### a) Univariate Analysis

```{r}
# Distribution of Health Risk Score
ggplot(df, aes(x = Health_Risk_Score)) +
  geom_histogram(bins = 50, fill = "#2993ae", alpha = 0.5) +
  labs(title = "Distribution of Health Risk Score", x = "Health Risk Score", y = "Frequency")
```

Most of the population has health risk scores around 9.0, with a small subset showing elevated scores near 10.0.

### b) Bivariate and Multivariate Analysis</span><br>



```{r}
summary_stats <- df %>%
  group_by(City) %>%
  summarise(
    Avg_Dew_Point = mean(dew, na.rm = TRUE),
    Avg_Humidity = mean(humidity, na.rm = TRUE),
    Avg_Pressure = mean(pressure, na.rm = TRUE),
    Total_Precipitation = sum(precip, na.rm = TRUE),
    pressure_a = sum(pressure, na.rm = TRUE),
    .groups = 'drop'  # Drop grouping
  )

# Print summary statistics
print(summary_stats)

# Visualization

# 1. Average Dew Point by City
ggplot(summary_stats, aes(x = City, y = Avg_Dew_Point)) +
  geom_bar(stat = "identity", fill = "lightblue") +
  labs(title = "Average Dew Point by City", x = "City", y = "Average Dew Point") +
  theme_minimal()

# 2. Average Humidity by City
ggplot(summary_stats, aes(x = City, y = Avg_Humidity)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  labs(title = "Average Humidity by City", x = "City", y = "Average Humidity") +
  theme_minimal()

# 3. Total Precipitation by City
ggplot(summary_stats, aes(x = City, y = Total_Precipitation)) +
  geom_bar(stat = "identity", fill = "lightcoral") +
  labs(title = "Total Precipitation by City", x = "City", y = "Total Precipitation") +
  theme_minimal()

# 3. Pressue by City
ggplot(summary_stats, aes(x = City, y = pressure_a)) +
  geom_bar(stat = "identity", fill = "#909090") +
  labs(title = "Pressure by City", x = "City", y = "Pressure") +
  theme_minimal()
```


This analysis explores key climate metrics—average dew point, average humidity, total precipitation, and atmospheric pressure across several major U.S. cities. These insights into varying atmospheric conditions can aid in understanding urban air quality and its potential health impacts.

**Average Dew Point by City**  
The average dew point reflects humidity levels across the cities analyzed. Houston and San Diego experience the highest average dew points (around 60°F), indicating more humid conditions in these areas. Los Angeles, however, has the lowest dew point, suggesting a drier climate.

**Average Humidity by City**  
The humidity chart highlights varying levels of moisture content. Cities like San Diego and Philadelphia have high humidity levels, exceeding 60%, which points to significant moisture in the air. In contrast, San Antonio and Dallas show relatively lower humidity levels, indicating generally drier conditions.

**Total Precipitation by City**  
This analysis shows the total precipitation levels across cities. Los Angeles and New York City have substantially higher precipitation, reflecting wetter climates or specific weather patterns. In contrast, cities like Chicago and San Jose show negative values, which could indicate dry spells, measurement anomalies, or a focus on the net precipitation balance.

**Pressure by City**  
The atmospheric pressure plot reveals average pressure levels across cities. New York City and Philadelphia exhibit higher cumulative pressure values, suggesting stable atmospheric conditions. Lower pressure values in San Diego and Los Angeles may indicate greater variability in atmospheric conditions in these cities.

The following points provides a detailed breakdown of various metrics across cities:

- **Avg_Dew_Point**: Average dew point temperature, highest in Houston and San Diego.

- **Avg_Humidity**: Average humidity levels, with San Diego and Philadelphia exhibiting significant moisture content.

- **Avg_Pressure**: Average atmospheric pressure, with New York City and Philadelphia on the higher end.

- **Total_Precipitation**: Net precipitation, showing substantial totals in Los Angeles and New York City, while San Jose has a negative value.

- **Pressure_a**: Sum of atmospheric pressure readings for each city, with San Jose having the highest cumulative pressure.

This comparative analysis of climate metrics provides insights into the atmospheric variations across urban environments, which can impact air quality and public health. The data underscores the diverse climate profiles of U.S. cities and their implications for air quality management and health outcomes.



```{r}
#Correlation between every columns
constant_columns <- sapply(df, function(x) length(unique(x)) == 1)
df <- df[, !constant_columns] 

#correlation matrix
cor_matrix <- cor(df[, sapply(df, is.numeric)])

# Melt the correlation matrix for plotting
melted_cor_matrix <- melt(cor_matrix)

# Correlation heatmap
ggplot(melted_cor_matrix, aes(Var1, Var2, fill = value)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(low = "green", high = "violet", mid = "white", midpoint = 0, limit = c(-1, 1)) +
  labs(title = "Correlation Heatmap", x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),axis.text.y = element_text(vjust=0,hjust=0))
```


Visualize the strength and direction of relationships (correlations) between pairs of variables in a dataset.

**Correlation with Health Risk Score**

A positive correlation means that as one variable increases, the other also increases. Conversely, a negative correlation means that as one variable increases, the other decreases.

Top **positive correlations:**

1. **Severity Score**: Strong positive link.
2. **Heat Index**: Significant positive association.
3. **Temperature Range**: Moderate positive correlation.
4. **Windgust**: Mild positive correlation.

Top **negative correlations:**

1. **Visibility**: Strong negative association.
2. **Solarenergy**: Significant negative correlation.
3. **Dew Point**: Moderate negative link.
4. **Humidity**: Mild negative correlation.

Wind gusts have a positive impact on health risk, whereas temperature has a weaker, slightly negative relationship with health risk scores.Wind-related factors (gusts and speed) and humidity increase health risk scores, while factors like visibility, sunrise/sunset times, and moon phases reduce them.

These indicate key factors influencing health risk scores.



### c) Smart Questions

```{r}
paste("Is there a statistically significant difference in the Health Risk Score on weekends compared to weekdays?")
  
# Separate Health Risk Scores by weekends and weekdays
weekend_scores <- df$Health_Risk_Score[df$Is_Weekend == "True"]
weekday_scores <- df$Health_Risk_Score[df$Is_Weekend == "False"]

# Display hypothesis statements
cat("Hypothesis Statements:\n")
cat("Null Hypothesis (H0): There is no significant difference in the Health Risk Score between weekends and weekdays.\n")
cat("Alternative Hypothesis (H1): There is a significant difference in the Health Risk Score between weekends and weekdays.\n\n")

t_test_result <- t.test(weekend_scores, weekday_scores, alternative = "two.sided")

# Display the t-test results
print(t_test_result)

# Interpretation
cat("\nInterpretation:\n")
if(t_test_result$p.value < 0.05) {
  cat("Reject the null hypothesis: There is a statistically significant difference in the Health Risk Score between weekends and weekdays.\n")
} else {
  cat("Fail to reject the null hypothesis: There is no statistically significant difference in the Health Risk Score between weekends and weekdays.\n")
}
```


```{r}
boxplot(Health_Risk_Score ~ Is_Weekend, data = df,
        main = "Health Risk Score Comparison: Weekends vs. Weekdays",
        xlab = "Is Weekend",
        ylab = "Health Risk Score",
        col = c("lightblue", "lightgreen"),
        border = "#202020")
```


**How do health risk scores differ between weekdays and weekends?**

There is a significant difference in Health Risk Scores between weekends and weekdays. Scores are generally higher on weekends than on weekdays, indicating increased health risks during weekends


```{r}
# Boxplot of Health Risk Score by City
ggplot(df, aes(x = City, y = Health_Risk_Score, fill = City)) +
  geom_boxplot() +
  labs(title = "City-wise Health Risk Score", x = "City", y = "Health Risk Score") +
  theme(axis.text.x = element_text(angle =90,hjust=1))
```


**How do key meteorological factors vary across different cities?**

**Key Observations**

- San Jose shows the widest IQR, indicating greater variability in health risk scores.
- Philadelphia has the lowest median health risk score, while New York City has the highest.
- Outliers are present in all cities, with New York City showing the most extreme values.

```{r}
# Boxplot of Windgust by City
ggplot(df, aes(x = City, y = windgust, fill = City)) +
  geom_boxplot() +
  labs(title = "City-wise windgust", x = "City", y = "Windgust") +
  theme(axis.text.x = element_text(angle =90,hjust=1))
```


**Key Observations**

- San Jose exhibits the widest IQR, indicating the most variability in wind gust speeds among the cities.
- Philadelphia has the lowest median wind gust speed, while Chicago has a relatively high median.
- New York City shows a few outliers, indicating instances of unusually high wind gust speeds.

San Jose has the highest health risk score and wind gusts, while Philadelphia shows the lowest health risk score and Los Angeles the lowest wind gusts.


```{r}
paste("Does wind gust have a significant impact on the Health Risk Score?")

paste("Correlation between windgust and Health risk :",cor(df$windgust, df$Health_Risk_Score))
plot1<-ggplot(df, aes(x = windgust, y = Health_Risk_Score)) +
  geom_point(color = "#2993ae") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Wind Gusts vs Health Risk Score", 
       x = "Wind Gust", 
       y = "Health Risk Score") +
  theme_minimal()

# Calculate correlation
paste("Correlation between severity score and Health risk :",cor(df$Severity_Score, df$Health_Risk_Score))
plot2<-ggplot(df, aes(x = Severity_Score, y = Health_Risk_Score)) +
  geom_point(color = "#2993ae") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Severity Score vs Health Risk Score", 
       x = "Severity Score", 
       y = "Health Risk Score") +
  theme_minimal()

paste("Correlation between wind speed and Health risk :",cor(df$windspeed, df$Health_Risk_Score))
plot3<-ggplot(df, aes(x = windspeed, y = Health_Risk_Score)) +
  geom_point(color = "#2993ae") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Wind speed vs Health Risk Score", 
       x = "Wind Speed", 
       y = "Health Risk Score") +
  theme_minimal()

paste("Correlation between humidity and Health risk :",cor(df$humidity, df$Health_Risk_Score))
plot4<-ggplot(df, aes(x = humidity, y = Health_Risk_Score)) +
  geom_point(color = "#2993ae") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "humidity vs Health Risk Score", 
       x = "humidity", 
       y = "Health Risk Score") +
  theme_minimal()

paste("Correlation between humidity and windgust :",cor(df$humidity, df$Health_Risk_Score))
plot5<-ggplot(df, aes(x = humidity, y = windgust)) +
  geom_point(color = "violet") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "humidity vs windgust", 
       x = "humidity", 
       y = "windgust") +
  theme_minimal()

grid.arrange(plot1, plot2,plot3,plot4,plot5, ncol =2 )
```


**Which key meteorological factors significantly influence health risk scores?**

**Direct Relationships with Health Risk Score:**

- **Wind Gusts**: The scatter plot shows a slight positive trend, suggesting that higher wind gusts are associated with a marginal increase in health risk scores.
- **Severity Score**: A strong positive correlation is observed, indicating that an increase in severity score corresponds to a notable rise in health risk scores.
- **Wind Speed**: Wind speed also displays a positive relationship with health risk score, implying that as wind speed increases, so does the health risk.
- **Humidity**: There is a positive association between humidity levels and health risk scores, with higher humidity associated with increased health risks.
    
**How might changes in humidity affect overall health?**

The analysis explores the relationships between environmental factors—such as wind gusts, severity score, wind speed, and humidity—and health risk scores.

**Indirect Influence of Humidity**:

- **Humidity vs. Wind Gust**: The scatter plot shows a slight positive correlation, suggesting that higher humidity may contribute to increased wind gusts. Since wind gusts are directly linked with higher health risk scores, this indicates that humidity may indirectly influence health risks through its effect on wind conditions.
    
        
```{r}
paste("Correlation between visibility and Health risk :",cor(df$windgust, df$Severity_Score))
plot1<-ggplot(df, aes(x = visibility, y = Health_Risk_Score)) +
  geom_point(color = "#287743") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Visibility vs Health risk score", 
       x = "visibility", 
       y = "Health_Risk_Score") +
  theme_minimal()

# Calculate correlation
paste("Correlation between sunsetEpoch and Health risk :",cor(df$Severity_Score, df$Health_Risk_Score))
plot2<-ggplot(df, aes(x = sunsetEpoch, y = Health_Risk_Score)) +
  geom_point(color = "#287743") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Sunset Epoch vs Health Risk Score", 
       x = "Sunset Epoch", 
       y = "Health Risk Score") +
  theme_minimal()

paste("Correlation between sunriseEpoch and Health risk :",cor(df$windspeed, df$Health_Risk_Score))
plot3<-ggplot(df, aes(x = sunriseEpoch, y = Health_Risk_Score)) +
  geom_point(color = "#287743") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Sunrise Epoch vs Health Risk Score", 
       x = "Sunrise Epoch", 
       y = "Health Risk Score") +
  theme_minimal()

paste("Correlation between moonphase and Health risk :",cor(df$windspeed, df$Health_Risk_Score))
plot4<-ggplot(df, aes(x = moonphase, y = Health_Risk_Score)) +
  geom_point(color = "#287743") +
  geom_smooth(method = "lm", color = "black") + # Adds a regression line
  labs(title = "Moon phase vs Health Risk Score", 
       x = "Moon phase", 
       y = "Health Risk Score") +
  theme_minimal()

grid.arrange(plot1, plot2,plot3,plot4, ncol = 2)
```

This analysis examines the relationship between certain environmental and astronomical factors—visibility, sunset epoch, sunrise epoch, and moon phase—and health risk scores.


- **Visibility**: The plot shows a slight negative correlation between visibility and health risk score. Higher visibility tends to be associated with lower health risks.
- **Sunset Epoch**: There is a negative trend between sunset time and health risk score, suggesting that later sunsets may be linked to reduced health risks.
- **Sunrise Epoch**: Similarly, a negative trend is observed between sunrise time and health risk score, with later sunrises associated with lower health risks.
- **Moon Phase**: The moon phase also shows a negative correlation with health risk score. As the moon progresses through its cycle, the health risk score tends to decrease.


These environmental and astronomical factors may subtly influence health risk scores. Clearer atmospheric conditions, later times of sunrise and sunset, and certain phases of the moon are associated with reduced health risks, suggesting possible connections between these natural phenomena and health outcomes.


```{r}
paste("How does the windgust vary over dates?")
  
# Convert datetime column to Date type (if not already done)
df$datetime <- as.Date(df$datetime, format = "%d-%m-%Y")  # Adjust the format as needed

# Extract month and day
df$Month <- format(df$datetime, "%m")  # Extract month as numeric
df$Day <- format(df$datetime, "%d")  # Extract day
# Average Health Risk Score by Date
daily_summary <- df %>%
  group_by(datetime) %>%
  summarise(Average_HRS = mean(windgust, na.rm = TRUE))

print(head(daily_summary))  # Preview the summary
library(ggplot2)

ggplot(daily_summary, aes(x = datetime, y = Average_HRS)) +
  geom_line( color = "blue") +
  geom_point(size = 1, color = "red") +
  labs(title = "Daily Windgust Score Over Time", 
       x = "Date", 
       y = "Windgust") +
  scale_x_date(date_breaks = "2 day", date_labels = "%d %b")
```


```{r}
paste("How does the Health Risk Score vary over dates?")
  
# Convert datetime column to Date type (if not already done)
df$datetime <- as.Date(df$datetime, format = "%d-%m-%Y")  # Adjust the format as needed

# Extract month and day
df$Month <- format(df$datetime, "%m")  # Extract month as numeric
df$Day <- format(df$datetime, "%d")  # Extract day
# Average Health Risk Score by Date
daily_summary <- df %>%
  group_by(datetime) %>%
  summarise(Average_HRS = mean(Health_Risk_Score, na.rm = TRUE))

print(head(daily_summary))  # Preview the summary
library(ggplot2)

ggplot(daily_summary, aes(x = datetime, y = Average_HRS)) +
  geom_line( color = "blue") +
  geom_point(size = 1, color = "red") +
  labs(title = "Daily Average Health Risk Score Over Time", 
       x = "Date", 
       y = "Average Health Risk Score") +
  scale_x_date(date_breaks = "2 day", date_labels = "%d %b")
```

**How does the windgust vary over dates?**

Health risk and wind gust scores both fluctuate similarly over time, peaking around September 8th and declining sharply by mid- September. Health risk score is being influenced by windgust.

Both graphs display a similar trend over the month of September. Initially, wind gusts are high, and health risk scores are elevated. As the month progresses, both the wind gust intensity and health risk scores decrease. By the end of the month, these values stabilize at lower levels. This refers to the correlation where reduced wind gusts are associated with lower health risks.

```{r}
daily_summary <- df %>%
  group_by(City, datetime) %>%  # Group by city and date
  summarise(Average_HRS = mean(Health_Risk_Score, na.rm = TRUE))

print(head(daily_summary))  # Preview the summary

# Load ggplot2 if not already loaded
library(ggplot2)

# Plot average Health Risk Score for each city
ggplot(daily_summary, aes(x = datetime, y = Average_HRS, color = City)) +
  geom_line() +  # Line for average scores
  geom_point(size = 1) +  # Points for average scores
  labs(title = "Health Risk Score Over Time by City", 
       x = "Date", 
       y = "Average Health Risk Score") +
  scale_x_date(date_breaks = "2 day", date_labels = "%d %b") +
  theme_minimal() +  # Use a minimal theme for better visualization
  theme(legend.position = "bottom")  # Place legend at the bottom
```

The above chart highlights variations in average health risk scores across U.S. cities during September. Some cities show steady trends, while others, like San Jose, experience significant daily fluctuations, suggesting a more dynamic health risk environment. Health scores generally range from 9.0 to 10.5, with some cities hitting higher levels early on, possibly indicating initial high-risk conditions that improved over time.

The dataset’s limitation to September, coupled with missing data for certain cities, restricts the ability to draw complete conclusions. In a real-time setting, missing data could hinder timely interventions. Future analyses should aim for consistent, long-term data to better capture health trends and support city planning efforts.


# **Conclusion and EDA Insights**

**Health Risk Variation:** Health risk scores are higher on weekends compared to weekdays, indicating greater risks during weekends

**City Comparison:** San Jose has the highest health risk and wind gusts, while Philadelphia shows the lowest health risk and wind gusts.

**Meteorological Impact:** Wind gusts significantly increase health risk scores, while temperature has a weaker influence.

**Humidity and Health:** Humidity was found to have a moderate positive correlation with health risk scores.

**Key Factors:** Wind-related factors increase health risks, while visibility and sunrise/sunset times lowers them indirectly.

# **Further Research**

**1. Standardize and Screen Health Risk Data Sources**  
To ensure robust and comparable analyses across locations, it is essential to establish standardized criteria for screening health risk data. This involves identifying reliable sources and uniform metrics for measuring health outcomes, such as respiratory or cardiovascular conditions influenced by air quality and meteorological factors. Expanding the research to include additional cities and diverse urban environments will enhance the study's generalizability.

**2. Expand the Research Scope with Time-Series Analyses**  
Conducting time-series analyses will allow for the exploration of how seasonal meteorological factors—like temperature, humidity, and precipitation—correlate with health risks over time. Time-series methods can identify trends, lags, and seasonal variations in health outcomes, offering a deeper understanding of how climate variability influences public health across different periods.

**3. Develop Mitigation Strategies for Extreme Weather-Related Health Impacts**  
Based on findings, targeted strategies can be developed to mitigate health risks associated with extreme weather conditions. This includes focusing on areas with high wind or humidity levels where the risk may be elevated. Potential strategies could involve community education, improvements in healthcare access during extreme weather events, and urban planning adjustments to reduce exposure to adverse conditions.

These directions will contribute to a more comprehensive understanding of urban air quality, weather patterns, and their health implications, ultimately supporting policy initiatives aimed at improving public health in urban areas.

# **References**

1. **Singh, R. K., Mishra, S. K., et al. (2022)**. *Urban Air Pollution, Urban Heat Island and Human Health: A Systematic Review*. *Sustainability*. [Link](https://www.mdpi.com/2071-1050/14/15/9234)

2. **Rojas, J. A. D., Weaver, C. P., et al. (2022)**. *Climate Change Impacts on Air Quality and Human Health*. *Data*. [Link](https://catalog.data.gov/dataset/climate-change-impacts-on-air-quality-and-human-health)

3. **Han, J., Zhang, W., et al. (2023)**. *Machine Learning for Urban Air Quality Analytics: A Survey*. [Link](https://arxiv.org/abs/2310.09620)










